Solving ...

### 2023-01-24 2023/PAVT_Ufa/PAVT2023.tex

\begin{document}
\mainmatter              % start of a contribution
% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
\title{Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm}
\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{
Irek Gubaydullin$^{1,2}$\and
Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
\and
Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
\and
Marina Usova$^4$\orcidID{XXX} 
}

@@

The paper considers the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number. As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed. To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters. To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data. From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem. A parallel information-statistical global search algorithm was used to solve it. The use of the parallel algorithm has significantly reduced the search time to find the optimum. The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.

@@

%Часть УГНТУ
Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number. Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons. The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline. Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью]. Therefore, it is a significant process of the modern refinery [ссылка на catalyst]. To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction. 
%Стыковочный абзац

Так как математической моделью реакции является система дифференциальных уравнений, найти найти значения констант, входящих в него, можно только численно (см., например, \cite{RSD}). Отметим, что в таких задачах целевая функция, как правило, является многоэкстремальной, т.е. имеет много локальных экстремумов наряду с глобальным. 
%Часть ННГУ
Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).  
С алгоритмической точки зрения их можно разделить на два класса: метаэвристические и детерминированные. Метаэвристические алгоритмы, в основном, основаны на имитации процессов, протекающих в живой природе.
Характерными примерами таких алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}). В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.  Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального \cite{Kvasov2018}.

>>>>

\begin{document}
\mainmatter              % start of a contribution
% Решение обратных задач химической кинетки с помощью асинхронного алгоритма глобального поиска
% Решение обратных задач химической кинетки с помощью смешанного локально-глобального поискового алгоритма 
%\title{Solving the Inverse Problems of Chemical Kinetics Using the Asynchronous Global Optimization Algorithm}
\title{
Solving ...
}
\titlerunning{Solving the Inverse Problems of Chemical Kinetics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{
Irek Gubaydullin$^{1,2}$\and
Leniza Enikeeva$^{2,3}$\orcidID{0000-0003-4219-4870}
\and
Konstantin Barkalov$^4$ \letter \orcidID{0000-0001-5273-2471}
\and
Marina Usova$^4$\orcidID{XXX} 
}

@@

The paper considers ...
%the application of parallel computing technology to the simulation of catalytic chemical reaction, which is widely used in the modern automobile industry to produce gasoline with high octane number. As a chemical reaction, the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid is assumed. To simulate a chemical process, it is necessary to develop a kinetic model of the process, that is, to determine the kinetic parameters. To do this, the inverse problem of chemical kinetics is solved, which predicts the values of kinetic parameters based on laboratory data. From a mathematical point of view, the inverse problem of chemical kinetics is a global optimization problem. A parallel information-statistical global search algorithm was used to solve it. The use of the parallel algorithm has significantly reduced the search time to find the optimum. The found optimal parameters of the model made it possible to adequately simulate the process of alkylation of isobutane with mixed C4 olefins catalyzed by sulfuric acid.

@@

%Часть УГНТУ
%Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number. Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons. The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline. Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью]. Therefore, it is a significant process of the modern refinery [ссылка на catalyst]. To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction. 
%Стыковочный абзац

Так как математической моделью реакции является система дифференциальных уравнений, найти значения констант, входящих в нее, можно только численно (см., например, \cite{RSD}). Отметим, что в таких задачах целевая функция, как правило, является многоэкстремальной, т.е. имеет много локальных экстремумов наряду с глобальным. 
%Часть ННГУ
Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}).  
С алгоритмической точки зрения их можно разделить на два класса: метаэвристические и детерминированные. Метаэвристические алгоритмы, в основном, основаны на имитации процессов, протекающих в живой природе.
Характерными примерами таких алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}). В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.  Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может располагаться далеко от глобального \cite{Kvasov2018}.
